---
title: "Skills & Qualifications for Statistics Jobs"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
```

# Hard Skills

## Column {data-width="350"}

### 

Highest Education Qualification Required

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#install.packages("devtools")
devtools::install_github("thiyangt/DSjobtracker")
library(DSjobtracker)
library(skimr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(scales)
library(ggradar)
library(patchwork)
drop <- c("ID","Consultant", "DateRetrieved", "DatePublished", "Company", "Payment Frequency", "Search_Term", "URL","Knowledge_in","City","Educational_qualifications")
DStidy_new <- DStidy[ , !(names(DStidy) %in% drop)]

```

```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width="150%"}
c1 <- sum(DStidy_new$BSc_needed==1 & DStidy_new$MSc_needed==0 & DStidy_new$PhD_needed==0,na.rm = TRUE)
c2 <- sum(DStidy_new$BSc_needed==0 & DStidy_new$MSc_needed==1 & DStidy_new$PhD_needed==0,na.rm = TRUE)
c3 <- sum(DStidy_new$BSc_needed==0 & DStidy_new$MSc_needed==0 & DStidy_new$PhD_needed==1,na.rm = TRUE)

c4 <- sum(DStidy_new$BSc_needed==1 & DStidy_new$MSc_needed==1 & DStidy_new$PhD_needed==0,na.rm = TRUE)

c5 <- sum(DStidy_new$BSc_needed==1 & DStidy_new$MSc_needed==0 & DStidy_new$PhD_needed==1,na.rm = TRUE)
c6 <- sum(DStidy_new$BSc_needed==0 & DStidy_new$MSc_needed==1 & DStidy_new$PhD_needed==1,na.rm = TRUE)
c7 <- sum(DStidy_new$BSc_needed==1 & DStidy_new$MSc_needed==1 & DStidy_new$PhD_needed==1,na.rm = TRUE)

cnt <- c(c1,c2+c4,c3+c5+c6+c7)
category <- c("BSc", "MSc","PhD")

df1 <- data.frame(category, cnt)

df1 <- df1  %>%
  mutate(pct = cnt / sum(cnt),
         pctlabel = paste0(round(pct*100), "%"))


ggplot(df1, 
       aes(x = reorder(category,cnt),
           y = cnt)) + 
  geom_bar(stat = "identity", 
           fill = "indianred3", 
           color = "black") +
  geom_text(aes(label = pctlabel), 
            vjust = -0.25) +
  labs(x = "Category", 
       y = "Count")  +
  coord_flip()+ theme(axis.title.x = element_blank())
```

### Highest Education Qualification based of the job Category

```{r, echo=FALSE, message=FALSE, warning=FALSE}
d1 <- DStidy_new  %>% filter(BSc_needed %in% 1)  %>% group_by(Job_Category) %>% count(BSc_needed)
d1 <- d1[,-2]
colnames(d1) <- c("Category", "BSc")
#View(d1)

d2 <- DStidy_new  %>% filter(MSc_needed %in% 1)  %>% group_by(Job_Category) %>% count(MSc_needed)
d2 <- d2[,-2]
colnames(d2) <- c("Category", "MSc")
#View(d2)

d3 <- DStidy_new  %>% filter(PhD_needed %in% 1)  %>% group_by(Job_Category) %>% count(PhD_needed)
d3 <- d3[,-2]
colnames(d3) <- c("Category", "PhD")
#View(d3)

df2 <- merge(d1, d2, by ="Category")
df2 <- merge(df2, d3, by ="Category", all = TRUE)
#View(df2)

df2 <- df2[-c(4,5),]
df2[is.na(df2)] <- 0


ggradar(
  df2, 
  values.radar = c("0", "20", "55"),
  grid.min = 0, grid.mid = 20, grid.max = 55,
  # Polygons
  group.line.width = 1, 
  group.point.size = 3,
  # Background and grid lines
  background.circle.colour = "white",
  gridline.mid.colour = "grey"
  ) + theme(legend.position = "bottom")

```

## Column {data-width="350"}

### Statistical Software Skills

```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width="150%"}
 
df5 <- DStidy_new  %>% filter(year %in% 2020)

ss1 <- sum(df5$`Ms Excel`==1 & df5$R ==0 & df5$SPSS ==0 & df5$Python ==0,na.rm = TRUE)
ss2 <- sum(df5$`Ms Excel`==0 & df5$R ==0 & df5$SPSS ==1 & df5$Python ==0,na.rm = TRUE)
ss3 <- sum(df5$`Ms Excel`==0 & df5$R ==1 & df5$SPSS ==0 & df5$Python ==0,na.rm = TRUE)
ss4 <- sum(df5$`Ms Excel`==0 & df5$R ==0 & df5$SPSS ==0 & df5$Python ==1,na.rm = TRUE)

ss5 <- sum(df5$R ==1 & df5$SPSS ==0 & df5$Python ==1,na.rm = TRUE)
ss6 <- sum(df5$R ==1 & df5$SPSS ==1 & df5$Python ==1,na.rm = TRUE)

df6 <- DStidy_new  %>% filter(year %in% 2021) 

ss11 <- sum(df6$`Ms Excel`==1 & df6$R ==0 & df6$SPSS ==0 & df6$Python ==0,na.rm = TRUE)
ss22 <- sum(df6$`Ms Excel`==0 & df6$R ==0 & df6$SPSS ==1 & df6$Python ==0,na.rm = TRUE)
ss33 <- sum(df6$`Ms Excel`==0 & df6$R ==1 & df6$SPSS ==0 & df6$Python ==0,na.rm = TRUE)
ss44 <- sum(df6$`Ms Excel`==0 & df6$R ==0 & df6$SPSS ==0 & df6$Python ==1,na.rm = TRUE)

ss55 <- sum(df6$R ==1 & df6$SPSS ==0 & df6$Python ==1,na.rm = TRUE)
ss66 <- sum(df6$R ==1 & df6$SPSS ==1 & df6$Python ==1,na.rm = TRUE)

Year_2020 <- c(ss1,ss2,ss3,ss4,ss5,ss6)
Year_2021 <- c(ss11,ss22,ss33,ss44,ss55,ss66)
Statistical_software <- c("Excel Only", "SPSS Only","R Only","Python Only","R & Python","R,Python & SPSS")

df7 <- data.frame(Statistical_software, Year_2020,Year_2021)
#View(df7)

df7 = df7 %>%
    gather(key=year, value=value, -Statistical_software)
#View(df7)

ggplot(data=df7, aes(x=Statistical_software, y=value, fill=Statistical_software)) +
    facet_wrap(~year) +
    geom_bar(stat='identity', position='dodge')+ 
    theme(legend.position = "none")  +
  theme(axis.text.x = element_text(angle = 45)) + theme(axis.title.x = element_blank())

```

### Programming Software Skills

```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width="150%"}
 
ps1 <- sum(df5$Java==1 & df5$Javascript ==0 & df5$Spark ==0,na.rm = TRUE)
ps2 <- sum(df5$Java==0 & df5$Javascript ==1 & df5$Spark ==0,na.rm = TRUE)
ps3 <- sum(df5$Java==0 & df5$Javascript ==0 & df5$Spark ==1,na.rm = TRUE)


ps11 <- sum(df6$Java==1 & df6$Javascript ==0 & df6$Spark ==0,na.rm = TRUE)
ps22 <- sum(df6$Java==0 & df6$Javascript ==1 & df6$Spark ==0,na.rm = TRUE)
ps33 <- sum(df6$Java==0 & df6$Javascript ==0 & df6$Spark ==1,na.rm = TRUE)


Year_2020 <- c(ps1,ps2,ps3)
Year_2021 <- c(ps11,ps22,ps33)
Programming_software <- c("Java Only", "JavaScript Only","Spark Only")

df8 <- data.frame(Programming_software, Year_2020,Year_2021)
#View(df7)

df8 = df8 %>%
    gather(key=year, value=value, -Programming_software)
#View(df7)

ggplot(data=df8, aes(x=Programming_software, y=value, fill=Programming_software)) +
    facet_wrap(~year) +
    geom_bar(stat='identity', position='dodge')+ 
    theme(legend.position = "none")  +
  theme(axis.text.x = element_text(angle = 45)) + theme(axis.title.x = element_blank())

```

# Soft Skills

## Column {data-width="350"}

### English Language Skills

```{r, echo=FALSE, message=FALSE, warning=FALSE}
 
e1 <- sum(DStidy_new$`English Needed`==1,na.rm = TRUE)

e2 <- sum(DStidy_new$`English Needed`==0,na.rm = TRUE)

cnt_e <- c(e1,e2)
english <- c("Needed", "Not Needed")

df3 <- data.frame(english, cnt_e )

df3 <- df3  %>%
  mutate(pct = cnt_e  / sum(cnt_e ),
         pctlabel = paste0(round(pct*100), "%"))


ggplot(df3, 
       aes(x = reorder(english, cnt_e),
           y = cnt_e)) + 
  geom_bar(stat = "identity", 
           fill = "indianred3", 
           color = "black") +
  geom_text(aes(label = pctlabel), 
            vjust = -0.25) +
  labs(x = "English", 
       y = "Count") 

```

## Column {data-width="350"}

### Soft Skills

```{r, echo=FALSE, message=FALSE, warning=FALSE}
 
s1 <- sum(DStidy_new$Presentation_Skills==1,na.rm = TRUE)

s2 <- sum(DStidy_new$Communication==1,na.rm = TRUE)

s3 <- sum(DStidy_new$Problem_Solving==1,na.rm = TRUE)

s4 <- sum(DStidy_new$Team_Handling==1,na.rm = TRUE)

cnt_s <- c(s1,s2,s3,s4)
soft <- c("Presentation", "Communication","Problem Solving","Team Building")

df4 <- data.frame(soft, cnt_s )

df4 <- df4  %>%
  mutate(pct = cnt_s  / sum(cnt_s),
         pctlabel = paste0(round(pct*100), "%"))


ggplot(df4, 
       aes(x = reorder(soft, cnt_s),
           y = cnt_s)) + 
  geom_bar(stat = "identity", 
           fill = "forestgreen", 
           color = "black") +
  geom_text(aes(label = pctlabel), 
            vjust = -0.25) +
  labs(x = "Soft Skill", 
       y = "Count") + coord_flip()

```

# Relationship between Salary and Skills

## Column {data-width="350"}

### Salary vs Education Qualifications

```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width="100%"}

df9 <- DStidy_new  %>% filter(BSc_needed %in% 1) 
df9 <- df9[df9$Job_Category != "Unimportant", ]
g1 <- ggplot(df9 %>% filter(!is.na(Job_Category)) , aes(x =Job_Category , y = Salary)) + geom_boxplot(outlier.shape = NA)  +
  theme(axis.text.x = element_text(angle = 45))  + 
  labs(y = "Salary")  +
  ggtitle("BSc Needed")+ theme(axis.title.x = element_blank()) +
  coord_cartesian(ylim =  c(0, 400000))


df10 <- DStidy_new  %>% filter(MSc_needed %in% 1) 
df10 <- df10[df10$Job_Category != "Unimportant", ]
g2 <- ggplot(df10 %>% filter(!is.na(Job_Category)) , aes(x =Job_Category , y = Salary)) + geom_boxplot(outlier.shape = NA)  +
  theme(axis.text.x = element_text(angle = 45)) + 
  labs(y = "Salary")  +
  ggtitle("MSc Needed") + theme(axis.title.x = element_blank()) +
  coord_cartesian(ylim =  c(0, 400000))


df11 <- DStidy_new  %>% filter(BSc_needed %in% 1) 
df11 <- df11[df11$Job_Category != "Unimportant", ]
g3 <- ggplot(df11 %>% filter(!is.na(Job_Category)) , aes(x =Job_Category , y = Salary)) + geom_boxplot(outlier.shape = NA)  +
  theme(axis.text.x = element_text(angle = 45)) + 
  labs(y = "Salary")  +
  ggtitle("PhD Needed")+ theme(axis.title.x = element_blank()) +
  coord_cartesian(ylim =  c(0, 400000))

ylab <- g1$labels$y
g1$labels$y <- g2$labels$y  <- g3$labels$y <- " "


g1+g2+g3
grid::grid.draw(grid::textGrob(ylab, x = 0.02, rot = 90))

```

## Column {data-width="350"}

### Salary vs Experience Level

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(DStidy_new %>% filter(!is.na(Experience_Category)) , aes(x =Experience_Category , y = Salary)) + geom_boxplot(outlier.shape = NA)  +
  theme(axis.text.x = element_text(angle = 45))  + 
  labs(x = "Experience level", y = "Salary")  +
  ggtitle("Relationship between Experience level & Salary") +
  coord_cartesian(ylim =  c(0, 2000000))

```
